---
title: "api_test"
format: html
---

## declare libraries and functions
```{r}
library(tidyverse)
library(arakno)
library(sf)
library(tmap)


get_mode <- function(x) {
  u <- unique(x)
  tab <- tabulate(match(x, u))
  u[tab == max(tab)]
}
```


## get all species in world spider catalog

### make list of families in dataset
```{r}
families = c("Agelenidae", "Anamidae", "Anapidae", "Anyphaenidae", "Araneidae", "Archaeidae", "Atracidae", "Atypidae", "Austrochilidae", "Barychelidae", "Bemmeridae", "Caponiidae", "Cheiracanthiidae", "Cithaeronidae", "Clubionidae", "Corinnidae", "Ctenidae", "Ctenizidae", "Cyatholipidae", "Cybaeidae", "Cycloctenidae", "Cyrtaucheniidae", "Deinopidae", "Desidae", "Dictynidae", "Diguetidae", "Drymusidae", "Dysderidae", "Entypesidae", "Eresidae", "Euagridae", "Euctenizidae", "Filistatidae", "Gallieniellidae", "Gnaphosidae", "Gradungulidae", "Hahniidae", "Halonoproctidae", "Hersiliidae", "Hexathelidae", "Homalonychidae", "Huttoniidae", "Hypochilidae", "Idiopidae", "Ischnothelidae", "Lamponidae", "Leptonetidae", "Linyphiidae","Linyphiidae","Linyphiidae","Linyphiidae", "Liocranidae", "Lycosidae", "Malkaridae", "Mecicobothriidae", "Mecysmaucheniidae", "Megadictynidae", "Microstigmatidae", "Migidae", "Mimetidae", "Miturgidae", "Mysmenidae", "Nemesiidae", "Nesticidae", "Ochyroceratidae", "Oecobiidae", "Oonopidae", "Orsolobidae", "Oxyopidae", "Pacullidae", "Palpimanidae", "Paratropididae", "Penestomidae", "Periegopidae", "Philodromidae", "Pholcidae", "Phrurolithidae", "Physoglenidae", "Phyxelididae", "Pisauridae", "Plectreuridae", "Psechridae", "Psilodercidae", "Pycnothelidae", "Scytodidae", "Segestriidae", "Selenopidae", "Sicariidae", "Sparassidae", "Stenochilidae", "Stiphidiidae", "Symphytognathidae", "Synaphridae", "Telemidae", "Tetrablemmidae", "Tetragnathidae", "Theraphosidae", "Theridiidae", "Theridiosomatidae", "Titanoecidae", "Toxopidae", "Trachelidae", "Trechaleidae", "Trochanteriidae", "Trogloraptoridae", "Udubidae", "Uloboridae", "Viridasiidae","Xenoctenidae", "Zodariidae", "Zoropsidae")
```

### ping the api and put all species into a df
```{r}
all_species <- data.frame(species(families)) %>% rename(species = species.families.) 
```

### fix or filter out species names that can't be read by api
```{r}
all_species_filt <- all_species %>% 
  mutate(
    
    # get word could for all latin names
    word_count = stringi::stri_count_words(species),
    
    # trim this one
    species = ifelse(
      species == 'Lycosa hispanica dufouri', 
      'Lycosa hispanica', species
      )
    ) %>% 
  
  # remove the rest of them that include some epithet beyond species
  filter(word_count < 3)
```

# get selected traits for every species in dataset
```{r}
print(cat("This code chunk pings the World Spider Catalog API,\nand fetches spider trait data for every species.\nIt takes hours to run, and you will overwrite the dataframe already stored."))
input = readline(prompt = "Continue? (yes/no): ")


if (input == "yes"){

  # make empty data frame and list to hold species that errored out
  all_traits <- data.frame(matrix(ncol = 48, nrow = 0))
  error_list <- list()
  
  # counter for printing every 200 species
  print_counter <- 1
  
  # we do tryCatch to allow species to error without disrupting the loop
  for (species in all_species_filt$species){
    tryCatch(
      {
        # increment the counter 
        print_counter <- print_counter + 1
        
        # and print the species every 200
        if (print_counter %% 200 == 0){
          print(paste("Current Index:", print_counter))
          print(paste("Current Species:", species))
        }
        
        # ping api for given species traits
        curr_spider_traits <- traits(species, 
                                     verbose = FALSE,
                                     trait = c(
                                       "bole", # body length
                                       "ld50", # venom toxicity
                                       "webb", # web building
                                       "retr", # retreat
                                       "cele", # cephalothorax length
                                       "cewe", # cephalothorax width
                                       "cehe", # cephalothorax height
                                       "fale", # fang length
                                       "vgsi", # venom gland size
                                       "strt", # stratum
                                       "circ" # circadian activity
                                        )
                                     )
        
        # skip to next species if nothing is returned
        if(length(curr_spider_traits) == 1 && curr_spider_traits == "No results found!") next
        
        # append to the df if something does come back
        all_traits = all_traits %>% rbind(curr_spider_traits)
      
      
      },
      
      # if error, print error and append said species to list
      error = function(e) {
        message(paste('Some error happened at this spider:', species))
        message(paste('Error in question:', e))
        
        error_list <<- append(error_list, species)
      }
    )
  }

  
  print("____execution completed____")
  
} else{
  print("____execution cancelled____")
}
```

```{r}

# read traits csv based off api ping above if necessary
if (!exists("all_traits")){
  all_traits <- read_csv("spider_traits.csv")
}

# a lot of data in world spider catalog requires authorization that I don't have
## which is why there are a lot of NAs
test <- all_traits %>% filter(!is.na(value))
print(nrow(all_traits)) # 71547 entries
nrow(test) # 45284 entries
#length(error_list) # 174 species omitted from dataset due to error
```



```{r}
print(cat("Overwrite the spider traits csv?"))
input = readline(prompt = "Continue? (yes/no): ")


if (input == "yes"){
  
  # save spider traits csv
  write.csv(all_traits, 'spider_traits.csv')

  print("____execution completed____")
  
} else{
  print("____execution cancelled____")
}
```



## Clean data spider data
```{r}
# pre filtering
all_traits_prefil <- all_traits %>% 
  
  # remove trait values with NAs (mostly restricted data)
  filter(!is.na(value)) %>%
  
  # select cols of interest
  select(id, originalName, value, sampleSize, note, rowLink, taxonomy.family, trait.abbrev, trait.name, measure.name, sex.name, country.name, location.coords.lat, location.coords.lon) %>% 
  
  # rename cols
  `colnames<-`(c('id', 'latin_name', 'value', 'sample_size', 'note', 'row_code', 'family', 'trait_abbrev', 'trait', 'measure_type', 'sex', 'country', 'lat', 'lon'))
```


## load countries and ecoregions data
```{r}
# load and fix polys
countries_shp <- read_sf("data/geoBoundariesCGAZ_ADM0/geoBoundariesCGAZ_ADM0.shp") %>% 
  st_make_valid() %>% 
  mutate(
    
    # calculate area
    country_area = st_area(geometry) %>%
      units::set_units(km^2)
    )

# load ecoregions (for biomes) and fix polygons
ecoregions <- read_sf("data/Terrestrial Ecoregions of the World/data/commondata/data0/wwf_terr_ecos.shp") %>% 
  st_make_valid()

# construct grid of points over the world map
point_grid <- SpatialPosition::CreateGrid(countries_shp, returnclass = "sf")
```


### clean and  filter ecoregions
```{r}
# match crs ecoregion crs with the country boundary shapefile
ecoregions <- st_transform(ecoregions, st_crs(countries_shp))

# union biomes together
ecoregions_union <- ecoregions %>% 
  
  # biomes have about 14 categories, which is not too much/little for our analysis
  group_by(BIOME) %>% 
  
  # union all biomes together
  summarize(biome_area = sum(AREA), geometry = st_union(geometry), 
            realm = get_mode(REALM_1)) %>%
  ungroup() %>%
  
  # include the full biome names in new col
  mutate(biome_nm = case_when(
          BIOME == 1 ~ "Tropical & Subtropical Moist Broadleaf Forests",
          BIOME == 2 ~ "Tropical & Subtropical Dry Broadleaf Forests",
          BIOME == 3 ~ "Tropical & Subtropical Coniferous Forests",
          BIOME == 4 ~ "Temperate Broadleaf & Mixed Forests",
          BIOME == 5 ~ "Temperate Conifer Forests",
          BIOME == 6 ~ "Boreal Forests/Taiga",
          BIOME == 7 ~ "Tropical & Subtropical Grasslands, Savannas & Shrublands",
          BIOME == 8 ~ "Temperate Grasslands, Savannas & Shrublands",
          BIOME == 9 ~ "Flooded Grasslands & Savannas",
          BIOME == 10 ~ "Montane Grasslands & Shrublands",
          BIOME == 11 ~ "Tundra",
          BIOME == 12 ~ "Mediterranean Forests, Woodlands & Scrub",
          BIOME == 13 ~ "Deserts & Xeric Shrublands",
          BIOME == 14 ~ "Mangroves",
          )
        )

# convert it to a df and get rid of undetermined biome vals
ecoregions_final <- ecoregions_union %>% 
  st_drop_geometry() %>% 
  filter(!BIOME %in% c(98,99))

# write out the csv
write_csv(ecoregions_final, "spider_ver_tables/ecoregions.csv")
```


### clean and filter country boundaries and export traits
```{r}

# case when object to align country names in country boundaries and world spider catalog datasets
country_case_when <- function(value){
  
  case_when(
  grepl("Congo", value) ~ "Congo", 
  grepl("Bolivia", value) ~ "Bolivia",
  grepl("Guiana", value) ~ "Guiana",
  grepl("Polynesia", value) ~ "Polynesia", 
  grepl("Iran", value) ~ "Iran", 
  value == "Korea, Republic of" | value == "Korea, South" ~ "South Korea", 
  value == "Lao People's Democratic Republic" ~ "Laos", 
  grepl("Moldova", value) ~ "Moldova", 
  grepl("Myanmar", value) ~ "Myanmar", 
  grepl("Macedonia", value) ~ "Macedonia",
  value == "Russian Federation" ~ "Russia",
  grepl("Sao Tome", value) ~ "Sao Tome & Principe",
  grepl("Solomon", value) ~ "Solomon Islands",
  grepl("Syria", value) ~ "Syria", 
  grepl("Taiwan", value) ~ "Taiwan", 
  grepl("Tanzania", value) ~ "Tanzania", 
  grepl("Trinidad", value) ~ "Trinidad & Tobago",
  grepl("United Kingdom", value) | grepl("Virgin Islands", value) ~ "United Kingdom", 
  grepl("United States", value) ~ "United States", 
  grepl("Venezuela", value) ~ "Venezuela", 
  grepl("Viet Nam", value) ~ "Vietnam",
  grepl("Bahamas", value) ~ "Bahamas",
  value == "Barbados" | value == "St Lucia" ~ "Bahamas",
  grepl("Gambia", value) ~ "Gambia", 
  grepl("d'Ivoire", value) ~ "Ivory Coast",
  grepl("Gaza Strip", value) ~ "Israel",
  grepl("Bosnia", value) ~ "Bosnia and Herzegovina",
  grepl("Guinea", value) ~ "Guinea",
  .default = value
  )
}
```


```{r}
all_traits_final <- all_traits_prefil %>% 
  
  # apply country name fix to traits dataset
  mutate(country = country_case_when(country)) %>% 
  
  # filter out any NAs from it and unecessary column
  filter(!is.na(country)) %>% 
  select(-trait_abbrev)

# do the same with the country boundaries
countries_shp_cor <- countries_shp %>% 
  mutate(shapeName = country_case_when(shapeName)) %>% 
  filter(!is.na(shapeName))


# # get list of all countries (that can join)
# all_countries <- all_traits_final %>% distinct(country) %>% inner_join(countries_shp_cor %>% distinct(shapeName), by = join_by(country == shapeName))
```


### join the countries to the ecoregions using the point grid
We use this method to get a more accurate calculation of how much of a particular ecoregion is in a given country
```{r}
# join points to biome
ecoregion_points <- point_grid %>% 
  st_join(ecoregions_union) %>% 
  filter(!is.na(BIOME))

# then join the pointy biomes to the countries
country_biome_points <- countries_shp_cor %>% 
  st_join(ecoregion_points)
```

```{r}
# get total points per biome
ecoregion_points_stats <- ecoregion_points %>% 
  count(BIOME) %>% # count points
  na.omit(n) %>% # omit biomes with 0 points
  as_tibble()


# calculate points per biome for each country, attach total biome points, calculate area of biome per country
country_eco_pts_stats <- country_biome_points %>% 
  
  # for each country in each biome
  group_by(shapeName, BIOME) %>% 
  
  # get count of points associated with each biome, for each country
  mutate(point_cnt = n()) %>% ungroup() %>% 
  
  # join the actual ecoregions dataset, to join area of biomes
  left_join(ecoregion_points_stats, by = join_by(BIOME)) %>% 
  
  # calculate area of biomes in each country by numer of points per biome
  mutate(biome_area_country = biome_area * (point_cnt/n)) %>%
  
  # convert to df
  st_drop_geometry() #%>% 
  # group_by(shapeName, BIOME, biome_area_country) %>% summarize() %>% ungroup() %>% 
  # rename(country = shapeName, biome = BIOME)
```


### join remaining countries
```{r}
# not all countries are joined with this method, so we try here to join some of them directly
## get all countries with NA biomes
strays <- country_eco_pts_stats %>% 
  filter(is.na(BIOME)) %>% 
  select(shapeName)


strays_attach <- strays %>% 
  
  # attach these to the country boundaries and make it geospatial
  inner_join(countries_shp, by = join_by(shapeName)) %>% 
  st_as_sf() %>% 
  
  # then join them with a simple polygon on polygon intersect
  st_join(ecoregions_union) %>% 
  
  # to df and clean
  st_drop_geometry() %>% 
  select(shapeName, country_area, BIOME, biome_area)


# then attach the stray countries to the df with the calculated stats
pre_countries_table <- country_eco_pts_stats %>% 
  
  # attach countries by name
  full_join(strays_attach, by = c("shapeName")) %>% 
  
  # fill in the NAs in the main biome cols
  mutate(BIOME.x = ifelse(is.na(BIOME.x), BIOME.y, BIOME.x), 
         biome_area.x = ifelse(is.na(biome_area.x), biome_area.y, biome_area.x)) %>% 
  
  # for rows containing the strays, grab the biome area from other rows
  group_by(BIOME.x) %>%
  mutate(n = ifelse(is.na(n), first(na.omit(n)), n)) %>%
  ungroup() %>%
  
  # calculate the areas for each biome within countries
  mutate(biome_area_country = ifelse(is.na(biome_area_country), biome_area.x * (point_cnt/n), biome_area_country)) %>%

  # df and cut down to grouped columns
  st_drop_geometry() %>%
  group_by(shapeName, BIOME.x, biome_area_country) %>% 
  summarize() %>% 
  ungroup() %>%
  
  # rename and regroup
  rename(country = shapeName, biome = BIOME.x)

pre_countries_table 
```


### create final countries table
```{r}
# countries table to be saved

final_countries_table <- pre_countries_table %>% 
  
  # for each country
  group_by(country) %>% 
  summarize(
    
    # grab the most dominant biome and it's area
    most_dominant_biome = biome[which.max(biome_area_country)], 
    most_dominant_biome_area = biome_area_country[which.max(biome_area_country)], 
    
    # and do the same thing for the second and third most dominant biomes
    second_dominant_biome = dplyr::nth(biome[order(-biome_area_country)], 2, default = NA),
    second_dominant_biome_area = dplyr::nth(biome_area_country[order(-biome_area_country)], 2, default = NA),
    third_dominant_biome = dplyr::nth(biome[order(-biome_area_country)], 3, default = NA),
    third_dominant_biome_area = dplyr::nth(biome_area_country[order(-biome_area_country)], 3, default = NA)
  )
final_countries_table
```


## Write out country boundaries and spider species traits datasets
```{r}
write_csv(final_countries_table, "spider_ver_tables/countries.csv")
write_csv(all_traits_final, "spider_ver_tables/full_spider_traits.csv")
```

